Implementing Capon Adaptive Beamforming on the GPU for Real Time Ultrasound Imaging

### 2012-05-04 SiPS2012/sips2012.tex
la2pdf commit - 2012:04:26	15:36:11	(Thu26.Apr)
Update in sips notes

% Title.
% ------
\title{Implementing Adaptive Beamforming on the GPU for Real Time Ultrasound Imaging}
%
% Single address.
% ---------------
\name{J. P. \AA{}sen$^1$, J. I. Buskenes$^2$, C.-I. C. Nilsen$^2$, A. Austeng$^2$ and S. Holm$^{1,2}$}
\address{$^1$Mi Lab, Norwegian University of Science and Technology, Trondheim, Norway \\ $^2$Department of Informatics, University of Oslo, Oslo, Norway}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}

@@

\section{Method}\label{sec:meth}
\subsection{Building the Sample Covariance Matrix}
Building $\hat{\mat{R}}$ using a sliding window approach across $K$ and $N_{avg}$. We have a limited amount of fast, near-core memory on the GPU. On NVIDIA architecture this memory is known as shared memory, and is restricted to 48KB per compute block. The maximum block size of 32x32, further restrict how large arrays we can handle inside a single compute block. It is therefore natural to restrict $L$ to a maxima of $32$ elements. Since $L \le M/2$, $M \le 64$. We can not afford to hold the full $\hat{\mat{R}}$ ($M = L$) when $\hat{\mat{R}} > 32$, because of limited amount of memory. ... Explain in detail how \texttt{buildR} is implemented in a kernel. ... Give more details on how much shared memory we have, and how it can be divided between compute blocks occupying one stream multiprocessor (SM).

@@

\subsubsection{CUDA Compute Model}
The CUDA compute model is based on execution of a kernel across a grid of compute threads. Each position in the grid holds a block of threads. Each block is further divided into groups of 32 threads, known as a warp. The maximum number of threads inside a block is restricted to 512, hence 16 warps. From a CUDA perspective the GPU consist of one or more stream multi processors (SM), where each SM is capable of 32 or more, depending on the architecture, concurrent multiplications and/or additions. The CUDA 2.x compute capability (CC) has available 48 KB of near-core shared memory per SM that can be utilized by resident blocks. The maximum number of resident blocks per SM is 8, but in addition the maximum number of resident warps per SM is restricted to 48. Hence, with a block size of 512 only $48/(512/32) = 3$ blocks can occupy the SM at once. In addition, if one thread where to use 64B of shared memory the number of allowed resident blocks with a block size of 512 would be reduced to 1. Up to a curtain point it is beneficial to have as many warps as possible occupying the SM at once to hide latency from accessing memory and executing time consuming mathematical functions. The SM has for CC 2.x 32K 32-bit long registers. When implementing algorithms on the GPU it is important to keep data as close to the core as possible. Utilizing available registers and shared memory to its maximum is therefore important.

\subsubsection{Optimizing kernel launch}

@@

Present images before and after Capon weights has been applied. Comment on the effect of selecting different parameters. Present graph showing running times for Cython-Capon v.s. CUDA-Capon for different choices of parameters.

@@

>>>>

% Title.
% ------
\title{Implementing Capon Adaptive Beamforming on the GPU for Real Time Ultrasound Imaging}
%
% Single address.
% ---------------
\name{J. P. \AA{}sen$^1$, J. I. Buskenes$^2$, C.-I. C. Nilsen$^2$, A. Austeng$^2$ and S. Holm$^{1,2}$}
\address{$^1$Mi Lab, Norwegian University of Science and Technology, Trondheim, Norway \\ $^2$Department of Informatics, University of Oslo, Oslo, Norway}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
%\twoauthors
%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
%	{School A-B\\
%	Department A-B\\
%	Address A-B}
%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
%	while at ...}}
%	{School C-D\\
%	Department C-D\\
%	Address C-D}
%
\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}

@@

\subsubsection{GPU Compute Model}
In this section we briefly discuss the principle behind GPU computing. CUDA by Nvidia has been our choice of GPU developer environment. However, even though CUDA has been used to implement the Capon beamformer, the challenges and proposed methods described in this paper will apply for other massive parallel implementation environments like e.g. OpenCL. The CUDA compute model is based on execution of a kernel across a grid of compute threads. Each position in the grid holds a block of threads. Each block is further divided into groups of 32 threads, known as a warp. The maximum number of threads inside a block is restricted to 512, hence 16 warps. From a CUDA perspective the GPU consist of one or more stream multi processors (SM), where each SM is capable of 32 or more, depending on the architecture, concurrent multiplications and/or additions. The CUDA 2.x compute capability (CC) has available 48 KB of near-core shared memory per SM that can be utilized by resident blocks. The maximum number of resident blocks per SM is 8, but in addition the maximum number of resident warps per SM is restricted to 48. Hence, with a block size of 512 only $48/(512/32) = 3$ blocks can occupy the SM at once. In addition, if one thread where to use 64B of shared memory the number of allowed resident blocks with a block size of 512 would be reduced to 1. Up to a curtain point it is beneficial to have as many warps as possible occupying the SM at once to hide latency from accessing memory. The SM has for CC 2.x 32K 32-bit long registers. When implementing algorithms on the GPU it is important to keep data as close to the core as possible. Utilizing available registers and shared memory to its maximum is therefore important.

\subsubsection{Optimizing kernel launch}
Set the size of grid adaptively based on $L$, $M$ and $Yavg$.

It is better to slide in space than time. $K$ is typically much larger than $2*Yavg + 1$.

\section{Parallel Capon}\label{sec:meth}
\subsection{Building Multiple Sample Covariance Matrices}
Building $\hat{\mat{R}}$ using a sliding window approach across $K$ and $N_{avg}$. We have a limited amount of fast, near-core memory on the GPU. On NVIDIA architecture this memory is known as shared memory, and is restricted to 48KB per compute block. The maximum block size of 32x32, further restrict how large arrays we can handle inside a single compute block. It is therefore natural to restrict $L$ to a maxima of $32$ elements. Since $L \le M/2$, $M \le 64$. We can not afford to hold the full $\hat{\mat{R}}$ ($M = L$) when $\hat{\mat{R}} > 32$, because of limited amount of memory. ... Explain in detail how \texttt{buildR} is implemented in a kernel. ... Give more details on how much shared memory we have, and how it can be divided between compute blocks occupying one stream multiprocessor (SM).

@@

@@

Present images before and after Capon weights has been applied. Comment on the effect of selecting different parameters. Present graph showing running times for Cython-Capon v.s. CUDA-Capon for different choices of parameters. Comment on different choices of parameters change the running times. Temporal averaging for instance, lowers the efficiency since $2*Y_{avg}$ more operations has to be performed per $\mat{\hat{R}}$

@@

\subsection{Compute Beamformer Output}
